---
title: "Assignment 1"
author: "Muhammad Uzair Aslam (18097) , Mustafa Qamar Shamim(17988),Sahad Morani(18000)"
date: "3/20/2021"
output: html_document
---

# Introduction
Our study is based on the monthly data of Exports, M2 and Consumer Price Index (CPI) of Pakistan. This data was collected from the archives of SBP bulletin.
The data from January 1992 till February 2021 was used to check whether these aforementioned time series are stationary or non-stationary. If non-stationary, the number of unit roots present in the time series were also calculated.
Testing data for stationarity is very important in research where the underlying variables based on time. If a series is non-stationary, it means that its variance or mean is dependent on time. With such a process, the effect of the change in the error term at any time will continue to the effect our series.
Stationarity shows that the statistical properties of the time series do not change over time. Thus, for our models and analysis to be reliable we need to make sure that the time series is stationary.

## Unit Root Tests

### DF Test
The Dickey–Fuller test tests the null hypothesis that a unit root is present in an autoregressive model. The alternative hypothesis is different depending on which version of the test is used, but is usually stationarity or trend-stationarity.

Suppose $Y_{t}$ is AR(1):
$$
Y_t=\phi Y_{t-1}+\varepsilon_t
$$
$$
Y_t-Y_{t-1}=\phi Y_{t-1}-Y_{t-1}+\varepsilon_t
$$
$$
Y_t-Y_{t-1}=(\phi-1)Y_{t-1}+\varepsilon_t
$$
$$
\Delta Y_t=\rho Y_{t-1}+\varepsilon_t
$$
With $\rho=\phi -1$


Where $\varepsilon_t \sim i.i.d.\mathcal{N}(0,\delta^2)$

$H_0: \rho=1$ There is a unit root

$H_1: \rho<1$ The series is stationary


### Augmented DF-Test
The Augmented Dickey-Fuller test allows for higher-order autoregressive processes in the model.

$$
Y_t=\beta_1+\beta_{2}t+ \phi Y_{t-1}+\varepsilon_t
$$
Which can be rewritten as:
$$
\Delta Y_t=\beta_1+\beta_{2}t+ \rho Y_{t-1}+\varepsilon_t
$$
Where $\rho=\phi -1$

$H_0: \rho=1$ There is a unit root present

$H_1: \rho <1$ This process is trend-stationary

### PP-Test

If the first differences of a series are serially correlated we should use the ADF test. However, detection of serial correlation along with appropriate selection lag length remain a difficult task. An alternative way to deal with this problem is the PP test

$Y_t=\phi Y_{t-1}+\varepsilon_t$, 

$\varepsilon_t \sim I(0)$ with mean $0$

Test Procedure:

1. Regress $Y_t$ on $Y_{t-1}$

2. Compute $\tau$

3. "Modify" $\tau$ to get z

4. Under $H_0$, z's asymptotic distribution is the DF distribution for $\tau$

### KPSS-Test
A series expressed as the sum of deterministic trend, random walk, and stationary error, is non-stationary.
The KPSS test is the LM test of the hypothesis that the random walk has zero variance. KPSS type tests are intended to complement unit root tests, such as the Dickey–Fuller tests.
Assume that we can decompose the series into the sum of a deterministic trend, a random walk and stationary errorwith the following linear regression model:
$$
Y_t=r_t+\beta T_t+ \varepsilon _t
$$
$H_0: \delta_{u}^2=0$ The series is stationary around a deterministic trend

$H_1: \delta_{u}^2>0$ The series is non-stationary

$H_0: \beta=0$ The series is stationary around a fixed level

$H_1: \beta>0$ The series is non-stationary

$r_t$ is a random walk process i.e. $r_t=r_{t-1}+u_t$

$u_t$ is iid $\mathcal{N}(0, \delta_{u}^2)$

$\beta T_t$ is a deterministic trend

$\varepsilon _t$ is a stationary error


### Zivot and Andrews Test

The problem with the ADF and PP test is that they don't allow for a structural break. Andrews and Zivot Test propose a variation of Perron’s original test in which they assume that the exact time of the break-point is unknown. Instead a data dependent algorithm is used to proxy Perron’s subjective procedure to determine the break points.
$$
\Delta x_t=\mu +\beta T_t+\lambda x_{t-1}+\phi_i \Delta x_{t-1}+\delta_1 DU_t+\delta_2 D(T)_t+\varepsilon_t
$$
Where,
DU is an indicator dummy variable for a mean shift occuring at each possible break-date

DT is corresponding trend shift variable

1. $DU = 1$ if $t>TB$ and $0$ otherwise

2. $Dt = t-TB$ if $t>TB$ and $0$ otherwise

$H_0: \alpha =0$ The series {${x_t}$} contains a unit root with a drift that excludes any structural break

$H_1: \alpha<0$ The series is a trend-stationary process with a one-time break occurring at an unknown point in time.  

### ERS test
$$
y_t=d_t+u_t
$$
$$
u_t=\alpha u_{t-1}+v_t
$$
Where {$d_t$} is a deterministic component and {$v_t$} is an unobserved stationary zero-mean error process whose spectral density function is positive at zero frequency.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
suppressWarnings({ 
library(ggplot2)
library(TSstudio)
library(forecast)
library(tseries)
library(Hmisc)
library(urca)
   })
```
```{r}
#clear global environment
rm(list = ls())

#below function reads the data from csv file
read_data <- function(){
        data <- read.csv("C:\\Users\\HP\\Downloads\\AE-II Assignment 1\\AEdata.csv")
        return(data)
}
data<- read_data()
describe(data)

```
```{r}
#below function converts the cpi into time series form
convert_cpi_to_ts <- function(){
        data <- read_data()
        cpi <- data$CPI
        cpi <- ts(cpi,frequency=12, start=c(1991,1))
        return(cpi)
}
```
```{r}
#below function converts the exports into time series form
convert_exports_to_ts <- function(){
        data <- read_data()
        exports <- data$Exports
        exports <- ts(exports,frequency=12, start=c(1991,1))
        return(exports)
}
```

```{r}
#below function converts the m2 into time series form
convert_m2_to_ts <- function(series){
        data <- read_data()
        m2 <- data$M2
        m2 <- ts(m2,frequency=12, start=c(1991,1))
        return(m2)
}
```
```{r}
#create instance of cpi and m2
cpi <- convert_cpi_to_ts()
m2  <- convert_m2_to_ts()
exports <- convert_exports_to_ts()
```

```{r}
#below function plots the cpi series in Time Series Studio
plot_cpi<- function(timeseries){
        #get cpi
        cpi <- convert_cpi_to_ts()
        
        #plot cpi
        p <-    ts_plot(cpi,
                title = "Line graph of CPI (January 1991- February 2021)",
                Xtitle = "Years",
                Ytitle = "CPI %",
                line.mode = "lines",
                slider = TRUE,
                Xgrid = TRUE,
                Ygrid = TRUE)
        return(p)
}
```
```{r}
plot_cpi()
```
The line graph of the CPI is increasing over time showing an increase in the prices of the goods. After 2015, inflation is increasing with a greater magnitude as compared to the previous years. 

```{r}
#below function plots the m2 series in Time Series Studio
plot_m2<- function(){
        #get m2
        m2 <- convert_m2_to_ts()
        
        #plot m2
        p <-    ts_plot(m2,
                        title = "Line graph of M2(January 1991- February 2021)",
                        Xtitle = "Years",
                        Ytitle = "M2",
                        line.mode = "lines",
                        slider = TRUE,
                        Xgrid = TRUE,
                        Ygrid = TRUE,
                        color = "Red")
        return(p)
}
```
```{r}
plot_m2()
```
The M2 money supply shows a constant increase with respect to time. This means that coins and currency in circulation,Savings deposits and Money Market funds have increased with time. 
```{r}
plot_exports<- function(){
        #get m2
        exports <-  convert_exports_to_ts()
        
        #plot m2
        p <-    ts_plot(exports,
                        title = "Line graph of Exports(January 1991- February 2021)",
                        Xtitle = "Years",
                        Ytitle = "Exports",
                        line.mode = "lines",
                        slider = TRUE,
                        Xgrid = TRUE,
                        Ygrid = TRUE,
                        color = "Green")
        return(p)
}
```
```{r}
plot_exports()
```
The exports show fluctuations between 1991 to 2020 but the general trend has been upwards. In the initial years up till 2002/2003, exports were relatively stable. From then on they have grown but with constant fluctuations that have also risen in magnitude.


```{r}
ts_decompose(cpi, type = "both")
```
Trend part in CPI is consistently upward and it does not satisfy mean reversion property because the mean CPI tends to change upward. Thus this is the non Stationary Series. Furthermore, there is a clear Seasonal trend in CPI. We can also see that in global crises there is a shock in Prices as CPI increased by 34.25% between 2007 and 2009. To further motivate,we made Seasonality plots to explore seasonal effects in CPI
```{r}
ts_decompose(m2, type = "both")
```
Trend part in M2 is consistently upwards, thus, it does not satisfy mean reversion property because the mean M2 tends to change upward. Therefore, this is concluded to be a non Stationary Series. Furthermore, there is a clear Seasonal trend in M2. 
```{r}
ts_decompose(exports, type = "both")
```
The trend of exports has a bit of variations but the overall trend is moving upwards, thus, it does not satisfy the mean diversion property. Therefore, this is concluded to be a non Stationary Series. Furthermore, there is a clear Seasonal trend in M2.
```{r}
ts_seasonal(cpi)
```
The plot reveals interesting patterns in CPI. Firstly we analyze that inflation has increased by more than double in last decade. In 2010 January Pakistan's CPI was 63.61 which rose to 132.77 in 2020 January. This is 108% increase in CPI between the decade.The rise of inflation has been fueled up after 2015 which can be seen from the plot.
```{r}
ts_seasonal(m2)
```

Comparing seasons over the years, growth in M2 has been exponential. January of 2020,reported 18.43m while at the start if the period in January 1990 were $347k. Throughout the time period, seasons of June July have shown the highest growth in M2 while October and November have been consistently stable.

```{r}
ts_seasonal(exports)
```

It is difficult to identify a consistent pattern in Pakistan's export over the period. Seasons have irregularly faced growth and fall signifying Pakistan's exports are unstable and do not depend much on seasonal factors.

DETERMINATION OF UNIT ROOTS
TECHNIQUE 1 : VISUALISING BY ACF AND PACF
Next we move on to find the acf and pacf of the series.The motivation is to find the stationarity of the series. So we plot them one by one

```{r}
ts_cor(cpi)
```
The ACF plot shows that CPI series is not Stationary since most of the spikes are out of 5% significant level. This means that mean,variance and covariances are time dependent.

AUGMENTED DF on CPI


```{r}
#we make a function which would take difference of the series two times
diff_2<- function(series)
{
         return(diff(diff(series)))
}
```

```{r}
#application of ADF test on cpi with trend and intercept
#H0: Ï = 1 there is a unit root
#H1: Ï < 1 the series is stationary
CPI_adf_level=adf.test(cpi)
CPI_adf_level
```
The P-value of the ADF test on level of CPI is 0.99 > 0.05. Thus, we accept the null hypothesis and conclude that this series is non-stationary at level.

```{r}
ndiffs(cpi)
```


```{r}
#now apply adf on second difference with trend and intercept
CPI_adf_diff2=adf.test(diff_2(cpi))
CPI_adf_diff2
```
The P-value of the ADF test on the second difference of CPI is 0.01 < 0.05. Thus, we reject the null hypothesis and conclude that this series is stationary at its second difference.


```{r}
#application of Dickey fuller on level with no trend and no intercept
CPI_df_type1=ur.df(cpi, type = c("none"), lags = 0, selectlags = c("Fixed"))
CPI_df_type1
```
The value of the test statistic is 14.3168 > -1.942 (Critical value of the ADF Test, model 0 at 5% significance). Thus, we accept the null hypothesis that this series is non-stationary.

```{r}
#application of Dickey fuller on level with trend and intercept
CPI_df_type3 <- urca::ur.df(cpi, type = "trend", lags = 0, selectlags = c("Fixed"))
urca::summary(CPI_df_type3)

```
The value of the test statistic is 1.1097 > -3.419 (Critical value of the ADF Test, model 2 at 5% significance). Thus, we accept the null hypothesis and conclude that this series is non-stationary.

```{r}
#application of Dickey fuller on second difference with trend and intercept
CPI_df_type3_diff2 <- urca::ur.df(diff_2(cpi), type = "trend", lags = 0, selectlags = c("Fixed"))
urca::summary(CPI_df_type3_diff2)

```
The value of the test statistic is -30.076 < -3.419 (Critical value of the ADF Test, model 2 at 5% significance). Thus, we reject the null hypothesis and conclude that this series is stationary.

```{r}
#application of Dickey fuller on level with intercept but no trend
CPI_df_type2 <- urca::ur.df(cpi, type = "drift", lags = 0, selectlags = c("Fixed"))
urca::summary(CPI_df_type2)

```
The value of the test statistic is 7.1529 > -2.873 (Critical value of the ADF Test, model 1 at 5% significance). Thus, we accept the null hypothesis and conclude that this series is non-stationary.

```{r}
#application of Augmented Dickey fuller on second difference
CPI_df_diff2 <- ur.df(diff_2(cpi), type = c("none"), lags = 0, selectlags = c("Fixed"))
CPI_df_diff2

```
The value of the test statistic is -30.1751 < -1.942 (Critical value of the ADF Test, model 0 at 5% significance). Thus, we reject the null hypothesis and conclude that this series is stationary.


```{r}
#applying KPSS on level
CPI_kpss_level <- kpss.test(cpi, null="Trend")
CPI_kpss_level
```
The P-value of this KPSS Test for Trend Stationarity is 0.01 < 0.05. Thus, we reject the null hypothesis and conclude that this series is non-stationary.

```{r}
#applying KPSS on difference 2
CPI_kpss_diff2 <- kpss.test(diff_2(cpi), null="Trend")
CPI_kpss_diff2

```
The P-value of this KPSS Test for Trend Stationarity is 0.1 > 0.05. Thus, we accept the null hypothesis and conclude that this series is trend-stationary.

```{r}
#applying ERS test for serial correlation on level
CPI_ers_level <- ur.ers(cpi, type="P-test", model="trend")
summary(CPI_ers_level)

```
For the Elliot, Rothenberg and Stock Unit Root Test, the value of the test statistic is 229.551 > 5.62 (at 5% level of significance). Thus, we reject the null hypothesis and conclude that serial correlation does not exist.

```{r}
#applying ERS test for serial correlation on difference 2
CPI_ers_diff2 <- ur.ers(diff_2(cpi), type="P-test", model="trend")
summary(CPI_ers_diff2)

```
For the Elliot, Rothenberg and Stock Unit Root Test, the value of the test statistic is 7.332 > 5.62 (at 5% level of significance). Thus, we reject the null hypothesis and conclude that serial correlation does not exist.


```{r}
#applying ELS to detrended data without intercept on level
CPI_ers_detrend <- ur.ers(cpi, type="DF-GLS", model="trend")
summary(CPI_ers_detrend)

```
For the Elliot, Rothenberg and Stock Unit Root Test, the value of the test statistic is 0.7866 > -2.89 (at 5% level of significance). Thus, we reject the null hypothesis and conclude that the series is stationary.


```{r}
#applying ELS to detrended data without intercept on difference
CPI_ers_detrend_diff1 <- ur.ers(diff(cpi), type="DF-GLS", model="trend")
summary(CPI_ers_detrend_diff1)

```
For the Elliot, Rothenberg and Stock Unit Root Test, the value of the test statistic is -5.7316 > -2.89 (at 5% level of significance). Thus, we accept the null hypothesis and conclude that the series is non-stationary.


```{r}
#applying Phillip and Perron test on level
CPI_pp_level <- pp.test(cpi)
CPI_pp_level

```
The P-value of Phillips-Perron Unit Root Test is 0.99 > 0.05. Thus, we accept the null hypothesis and conclude that the series is non-stationary.

```{r}
#applying Phillip and Perron test on difference 2
CPI_pp_diff2 <- pp.test(diff_2(cpi))
CPI_pp_diff2
```

The P-value of Phillips-Perron Unit Root Test is 0.01 < 0.05. Thus, we reject the null hypothesis and conclude that the series is stationary.


```{r}
#Performs the Zivot \& Andrews unit root test, which allows a break at an unknown point in either the intercept, the linear trend or in both.
#performing on level
# H0 :yt =Î¼+ytâ1 +Îµt
CPI_az_level <- ur.za(cpi , model="both", lag=2)
summary(CPI_az_level)
```
The test statistic value of the Zivot-Andrews Unit Root Test is -2.1708 > -5.08 (at 5% level of significance). This shows that the series is non-stationary at level and there is a potential break at position 119

```{r}
#plotting on level
plot(CPI_az_level)

```


```{r}
#Performs the Zivot \& Andrews unit root test, which allows a break at an unknown point in either the intercept, the linear trend or in both.
#performing on difference 2
# H0 :yt =Î¼+ytâ1 +Îµt
CPI_az_diff2 <- ur.za(diff_2(cpi) , model="both", lag=2)
summary(CPI_az_diff2)
```
The test statistic value of the Zivot-Andrews Unit Root Test is -16.7729 < -5.08 (at 5% level of significance). The test reports a rejection of the unit root null in favour of a one time break in the intercept at position 350.
```{r}
#at difference of 2 that is stationary
plot(CPI_az_diff2)


```


```{r}
ts_cor(diff_2(cpi))


```
The ACF plot shows that CPI series is Stationary at the second difference since most of the spikes are out of 5% significant level. This means that mean,variance and covariances are time independent.

```{r}
ts_cor(m2)
```
The ACF plot shows that M2 series is not Stationary at level since most of the spikes are out of 5% significant level. This means that mean,variance and covariances are time dependent.
```{r}
#application of ADF test on m2 with trend and intercept
#H0: Ï = 1 there is a unit root
#H1: Ï < 1 the series is stationary
m2_adf_level=adf.test(m2)
m2_adf_level
```
The P-value of the ADF test on level of M2 is 0.99 > 0.05. Thus, we accept the null hypothesis and conclude that this series is non-stationary at level.
```{r}
ndiffs(m2)


#now apply adf on second difference with trend and intercept
m2_adf_diff2=adf.test(diff_2(m2))
m2_adf_diff2
```
The P-value of the ADF test on second difference of M2 is 0.01 < 0.05. Thus, we reject the null hypothesis and conclude that this series is stationary at second difference.
```{r}
#application of Dickey fuller on level with no trend and no intercept
m2_df_type1=ur.df(m2, type = c("none"), lags = 0, selectlags = c("Fixed"))
m2_df_type1
```
The value of the test statistic is 11.1343 > -1.942 (Critical value of the ADF Test, model 0 at 5% significance). Thus, we accept the null hypothesis that this series is non-stationary.
```{r}
#application of Dickey fuller on level with trend and intercept
m2_df_type3 <- urca::ur.df(m2, type = "trend", lags = 0, selectlags = c("Fixed"))
urca::summary(m2_df_type3)
```
The value of the test statistic is 2.4153 > -3.42 (Critical value of the ADF Test, model 2 at 5% significance). Thus, we accept the null hypothesis and conclude that this series is non-stationary at level.

```{r}
#application of Dickey fuller on second difference with trend and intercept
m2_df_type3_diff2 <- urca::ur.df(diff_2(m2), type = "trend", lags = 0, selectlags = c("Fixed"))
urca::summary(m2_df_type3_diff2)
```
The value of the test statistic is -32.0662 < -3.42 (Critical value of the ADF Test, model 2 at 5% significance). Thus, we reject the null hypothesis and conclude that this series is stationary at second difference.

```{r}
#application of Dickey fuller on level with intercept but no trend
m2_df_type2 <- urca::ur.df(m2, type = "drift", lags = 0, selectlags = c("Fixed"))
urca::summary(m2_df_type2)
```
The value of the test statistic is 7.5375 > -2.873 (Critical value of the ADF Test, model 1 at 5% significance). Thus, we accept the null hypothesis and conclude that this series is non-stationary.
```{r}
#application of Augmented Dickey fuller on second difference
m2_df_diff2 <- ur.df(diff_2(m2), type = c("none"), lags = 0, selectlags = c("Fixed"))
m2_df_diff2
```
The value of the test statistic is -32.1544 < -1.942 (Critical value of the ADF Test, model 0 at 5% significance). Thus, we reject the null hypothesis and conclude that this series is stationary.
```{r}
#applying KPSS on level
m2_kpss_level <- kpss.test(m2, null="Trend")
m2_kpss_level
```
The P-value of this KPSS Test for Trend Stationarity is 0.01 < 0.05. Thus, we reject the null hypothesis and conclude that this series is non-stationary.

```{r}
#applying KPSS on difference 2
m2_kpss_diff2 <- kpss.test(diff_2(m2), null="Trend")
m2_kpss_diff2
```
The P-value of this KPSS Test for Trend Stationarity is 0.1 > 0.05. Thus, we accept the null hypothesis and conclude that this series is stationary.

```{r}
#applying ERS test for serial correlation on level
m2_ers_level <- ur.ers(m2, type="P-test", model="trend")
summary(m2_ers_level)
```
For the Elliot, Rothenberg and Stock Unit Root Test, the value of the test statistic is 2038.528 > 5.62 (at 5% level of significance). Thus, we reject the null hypothesis and conclude that serial correlation does not exist.

```{r}
#applying ERS test for serial correlation on difference 2
m2_ers_diff2 <- ur.ers(diff_2(m2), type="P-test", model="trend")
summary(m2_ers_diff2)
```
For the Elliot, Rothenberg and Stock Unit Root Test, the value of the test statistic is 160.6301 > 5.62 (at 5% level of significance). Thus, we reject the null hypothesis and conclude that serial correlation does not exist.

```{r}
#applying ELS to detrended data without intercept on level
m2_ers_detrend <- ur.ers(m2, type="DF-GLS", model="trend")
summary(m2_ers_detrend)
```
For the Elliot, Rothenberg and Stock Unit Root Test, the value of the test statistic is 1.9329 > -2.89 (at 5% level of significance). Thus, we reject the null hypothesis and conclude that the series is stationary.


```{r}
#applying ELS to detrended data without intercept on difference
m2_ers_detrend_diff1 <- ur.ers(diff(m2), type="DF-GLS", model="trend")
summary(m2_ers_detrend_diff1)
```
For the Elliot, Rothenberg and Stock Unit Root Test, the value of the test statistic is -10.7818 < -2.89 (at 5% level of significance). Thus, we can not reject the null hypothesis and conclude that the series is non-stationary.

```{r}
#applying Phillip and Perron test on level
m2_pp_level <- pp.test(m2)
m2_pp_level
```
The P-value of Phillips-Perron Unit Root Test is 0.99 > 0.05. Thus, we accept the null hypothesis and conclude that the series is non-stationary.
```{r}
#applying Phillip and Perron test on difference 2
m2_pp_diff2 <- pp.test(diff_2(m2))
m2_pp_diff2
```
The P-value of Phillips-Perron Unit Root Test is 0.01 < 0.05. Thus, we reject the null hypothesis and conclude that the series is stationary.
```{r}
#Performs the Zivot \& Andrews unit root test, which allows a break at an unknown point in either the intercept, the linear trend or in both.
#performing on level
# H0 :yt =Î¼+ytâ1 +Îµt
m2_az_level <- ur.za(m2 , model="both", lag=2)
summary(m2_az_level)
```
The test statistic value of the Zivot-Andrews Unit Root Test is 1.128 > -5.08 (at 5% level of significance). This shows that the series is non-stationary at level and there is a potential break at position 230
```{r}
#plotting on level
plot(m2_az_level)



#Performs the Zivot \& Andrews unit root test, which allows a break at an unknown point in either the intercept, the linear trend or in both.
#performing on difference 2
# H0 :yt =Î¼+ytâ1 +Îµt
m2_az_diff2 <- ur.za(diff_2(m2) , model="both", lag=2)
summary(m2_az_diff2)
```
The test statistic value of the Zivot-Andrews Unit Root Test is -16.9124 > -5.08 (at 5% level of significance). This shows that the series is stationary at second difference and there is a potential break at position 352

```{r}
#at difference of 2 that is stationary
plot(m2_az_diff2)



ts_cor(diff_2(m2))
```


```{r}
ts_cor(exports)
```
The ACF plot shows that Exports series is not Stationary at level since most of the spikes are out of 5% significant level. This means that mean,variance and covariances are time dependent.
```{r}
#application of ADF test on exports with trend and intercept
#H0: Ï = 1 there is a unit root
#H1: Ï < 1 the series is stationary
exports_adf_level=adf.test(exports)
exports_adf_level
```
The P-value of the ADF test on level of Exports is 0.343 > 0.05. Thus, we accept the null hypothesis and conclude that this series is non-stationary at level.

```{r}
ndiffs(exports)


#now apply adf on first difference with trend and intercept
exports_adf_diff2=adf.test(diff(exports))
exports_adf_diff2
```
The P-value of the ADF test on second difference of Exports is 0.01 < 0.05. Thus, we reject the null hypothesis and conclude that this series is stationary at first difference.
```{r}
#application of Dickey fuller on level with no trend and no intercept
exports_df_type1=ur.df(exports, type = c("none"), lags = 0, selectlags = c("Fixed"))
exports_df_type1
```
The value of the test statistic is -0.5803 > -1.942 (Critical value of the ADF Test, model 0 at 5% significance). Thus, we accept the null hypothesis that this series is non-stationary.
```{r}
#application of Dickey fuller on level with trend and intercept
exports_df_type3 <- urca::ur.df(exports, type = "trend", lags = 0, selectlags = c("Fixed"))
urca::summary(exports_df_type3)
```
The value of the test statistic is 6.5274 > -3.42 (Critical value of the ADF Test, model 2 at 5% significance). Thus, we accept the null hypothesis and conclude that this series is non-stationary at level.

```{r}
#application of Dickey fuller on first difference with trend and intercept
exports_df_type3_diff1 <- urca::ur.df(diff(exports), type = "trend", lags = 0, selectlags = c("Fixed"))
urca::summary(exports_df_type3_diff1)
```
The value of the test statistic is -31.3129 < -3.42 (Critical value of the ADF Test, model 2 at 5% significance). Thus, we reject the null hypothesis and conclude that this series is stationary at first difference.

```{r}
#application of Dickey fuller on level with intercept but no trend
exports_df_type2 <- urca::ur.df(exports, type = "drift", lags = 0, selectlags = c("Fixed"))
urca::summary(exports_df_type2)
```
The value of the test statistic is -2.4943 > -2.873 (Critical value of the ADF Test, model 1 at 5% significance). Thus, we accept the null hypothesis and conclude that this series is non-stationary.
```{r}
#application of Augmented Dickey fuller on first difference
exports_df_diff1 <- ur.df(diff(exports), type = c("none"), lags = 0, selectlags = c("Fixed"))
exports_df_diff1
```
The value of the test statistic is -31.9638 < -2.873 (Critical value of the ADF Test, model 1 at 5% significance). Thus, we reject the null hypothesis and conclude that this series is stationary at first difference.
```{r}
#applying KPSS on level
exports_kpss_level <- kpss.test(exports, null="Trend")
exports_kpss_level
```
The P-value of this KPSS Test for Trend Stationarity is 0.01 < 0.05. Thus, we reject the null hypothesis and conclude that this series is non-stationary.
```{r}
#applying KPSS on difference 1
exports_kpss_diff1 <- kpss.test(diff(exports), null="Trend")
exports_kpss_diff1
```
The P-value of this KPSS Test for Trend Stationarity is 0.1 > 0.05. Thus, we accept the null hypothesis and conclude that this series is stationary on first difference.
```{r}
#applying ERS test for serial correlation on level
exports_ers_level <- ur.ers(exports, type="P-test", model="trend")
summary(exports_ers_level)
```
For the Elliot, Rothenberg and Stock Unit Root Test, the value of the test statistic is 4.7791 < 5.62 (at 5% level of significance). Thus, we accept the null hypothesis and conclude that serial correlation does exist.
```{r}
#applying ERS test for serial correlation on difference 1
exports_ers_diff1 <- ur.ers(diff(exports), type="P-test", model="trend")
summary(exports_ers_diff1)
```
For the Elliot, Rothenberg and Stock Unit Root Test, the value of the test statistic is 0.003 < 5.62 (at 5% level of significance). Thus, we accept the null hypothesis and conclude that serial correlation does exist.

```{r}
#applying ELS to detrended data without intercept on level
exports_ers_detrend <- ur.ers(exports, type="DF-GLS", model="trend")
summary(exports_ers_detrend)
```
For the Elliot, Rothenberg and Stock Unit Root Test, the value of the test statistic is -3.3343 < -2.89 (at 5% level of significance). Thus, we accept the null hypothesis and conclude that the series is non-stationary on level.
```{r}
#applying ELS to detrended data without intercept on difference
exports_ers_detrend_diff1 <- ur.ers(diff(exports), type="DF-GLS", model="trend")
summary(exports_ers_detrend_diff1)
```
For the Elliot, Rothenberg and Stock Unit Root Test, the value of the test statistic is -8.7273 < -2.89 (at 5% level of significance). Thus, we accept the null hypothesis and conclude that the series is non-stationary on first difference.

```{r}
#applying Phillip and Perron test on level
exports_pp_level <- pp.test(exports)
exports_pp_level
```
The P-value of Phillips-Perron Unit Root Test is 0.01 < 0.05. Thus, we reject the null hypothesis and conclude that the series is stationary at level.
```{r}
#applying Phillip and Perron test on difference 1
exports_pp_diff1 <- pp.test(diff(exports))
exports_pp_diff1
```
The P-value of Phillips-Perron Unit Root Test is 0.01 < 0.05. Thus, we reject the null hypothesis and conclude that the series is stationary at first difference.
```{r}
#Performs the Zivot \& Andrews unit root test, which allows a break at an unknown point in either the intercept, the linear trend or in both.
#performing on level
# H0 :yt =Î¼+ytâ1 +Îµt
exports_az_level <- ur.za(exports , model="both", lag=2)
summary(exports_az_level)
```
The test statistic value of the Zivot-Andrews Unit Root Test is -4.6112 > -5.08 (at 5% level of significance). This shows that the series is non-stationary at level and there is a potential break at position 230
```{r}
#plotting on level
plot(exports_az_level)


#Performs the Zivot \& Andrews unit root test, which allows a break at an unknown point in either the intercept, the linear trend or in both.
#performing on difference 1
# H0 :yt =Î¼+ytâ1 +Îµt
exports_az_diff1 <- ur.za(diff(exports) , model="both", lag=2)
summary(exports_az_diff1)



#at difference of 2 that is stationary
plot(exports_az_diff1)
```
The test statistic value of the Zivot-Andrews Unit Root Test is -13.1821 < -5.08 (at 5% level of significance). This shows that the series is stationary at first difference and there is a potential break at position 347
```{r}
ts_cor(diff(exports))
